---
title: "Assignment 1: Group Shannon"
output: html_notebook
---

# Prithvi Poddar (17191) 

```{r}
library(ggplot2)
library(tidyverse)
library(dplyr)
library(tidyr)
library(Hmisc)
library(plyr)
library(readr)
library(GGally)
library(caret)
```

Setting the working directory
```{r}
setwd("~/Documents/iiser works/ECS/8th_sem/Bio_stats")
```

# Comparison among developed, developing and under developed countries
    
Here, we take 10 countries from each **Developed**, **Developing** and **Under Developed** countries. We then find the year-wise mean for the 10 countries and plot the values from the year **1800 to 2100**. Then we plot the graphs for these three groups and find out some patterns or test various hypothesis on them.    
The countries in the lists are:    
**Developed countries       :** Norway , Ireland , Switzerland , Hong Kong , Iceland , Germany , Sweden , Australia, Netherlands , Denmark    
**Developing countries      :** Algeria , Lebanon , Fiji , Moldova , Maldives , Tunisia , Saint Vincent and the Grenadines , Suriname , Mongolia , Botswana    
**Under Developed Countries :** Eritrea , Mozambique , Burkina Faso , Sierra Leone , Mali , Burundi , South Sudan , Chad , Central African Republic, Niger   
Above set of countries have been selected based on their Human Development Index rankings published in year 2020 by United Nations
Development Program (http://www.hdr.undp.org/).

**First we load all the datasets**
```{r}
read.csv("child_mortality_0_5_year_olds_dying_per_1000_born.csv", header = T, check.names = F) -> ChildMortality
read.csv("children_per_woman_total_fertility.csv", header = T, check.names = F) -> ChildrenPerWomen
read.csv("income_per_person_gdppercapita_ppp_inflation_adjusted.csv", header = T, check.names = F)-> IncomePerPerson
read.csv("life_expectancy_years.csv", header = T, check.names = F)-> LifeExpectancy
read.csv("population_total.csv", header = T, check.names = F)-> Population
```
    
Next, we extract the grouped data of the countries based on whether they are developed, developing or under developed. To do that, we first write a funtion as follows:   
```{r}
Get_country_data <- function(Dataset, a, b, c, d, e, f, g, h, i, j) {
  Dataset[Dataset$country %in% c(a, b, c, d, e, f, g, h, i, j), ]}
```
    
Getting the data for **Developed Countries**
```{r}
a<-"Norway"
b<-"Ireland"
c<-"Switzerland"
d<-"Finland"
e<-"Iceland"
f<-"Germany"
g<-"Sweden"
h<-"Australia"
i<-"Netherlands"
j<-"Denmark"

TopLife <- Get_country_data(LifeExpectancy, a, b, c, d, e, f, g, h, i, j)
TopPopulation <- Get_country_data(Population, a, b, c, d, e, f, g, h, i, j)
TopIncome <- Get_country_data(IncomePerPerson, a, b, c, d, e, f, g, h, i, j)
TopChildren <- Get_country_data(ChildrenPerWomen, a, b, c, d, e, f, g, h, i, j)
TopMortality <- Get_country_data(ChildMortality, a, b, c, d, e, f, g, h, i, j)
```
    
Getting the data for **Developing Countries**
```{r}
a<-"Algeria"
b<-"Lebanon"
c<-"Fiji"
d<-"Moldova"
e<-"Maldives"
f<-"Tunisia"
g<-"Saint Vincent and the Grenadines" 
h<-"Suriname"
i<-"Mongolia"
j<-"Botswana"

MidLife <- Get_country_data(LifeExpectancy, a, b, c, d, e, f, g, h, i, j)
MidPopulation <- Get_country_data(Population, a, b, c, d, e, f, g, h, i, j)
MidIncome <- Get_country_data(IncomePerPerson, a, b, c, d, e, f, g, h, i, j)
MidChildren <- Get_country_data(ChildrenPerWomen, a, b, c, d, e, f, g, h, i, j)
MidMortality <- Get_country_data(ChildMortality, a, b, c, d, e, f, g, h, i, j)
```
    
Getting the data for **Under Developed Countries**
```{r}
a<-"Eritrea"
b<-"Mozambique"
c<-"Burkina Faso"
d<-"Sierra Leone"
e<-"Mali"
f<-"Burundi"
g<-"South Sudan" 
h<-"Chad"
i<-"Central African Republic"
j<-"Niger"

LowLife <- Get_country_data(LifeExpectancy, a, b, c, d, e, f, g, h, i, j)
LowPopulation <- Get_country_data(Population, a, b, c, d, e, f, g, h, i, j)
LowIncome <- Get_country_data(IncomePerPerson, a, b, c, d, e, f, g, h, i, j)
LowChildren <- Get_country_data(ChildrenPerWomen, a, b, c, d, e, f, g, h, i, j)
LowMortality <- Get_country_data(ChildMortality, a, b, c, d, e, f, g, h, i, j)
```
    
    
## DataSet: Life Expentency 
**Finding the year wise mean for all the countries**    
```{r}
TopLifeMean <- sapply(TopLife[,2:302], mean)
MidLifeMean <- sapply(MidLife[,2:302], mean)
LowLifeMean <- sapply(LowLife[,2:302], mean)
```

**Plotting the means for all the countries for all the 3 groups in the same plot, to compare the the trends**
```{r}
plot(c(1800:2100), TopLifeMean, col="red", main = "Life Expentency in years", pch=1, ylim = c(0,150), xlab = "Years", ylab = "Life Expectency") 
points(c(1800:2100), MidLifeMean, col="blue")
points(c(1800:2100), LowLifeMean, pch=20, col="green")
legend(x = "topright", legend = c("Developed countries ","Developing countries","Under Developed countries"), fill = c("red","blue","green"))
```
    
**Making box plot to compare the means over the course of years**
```{r}
boxplot(TopLifeMean, MidLifeMean, LowLifeMean, ylab = "Life expectency in years", names=c("Developed ","Developing ","Under Developed "))
```
**Analysis:**   
In the initial years, the life expectancy for developing and under developed countries were the same with the developed countries slightly above them. From around 1860, the life expectency of developed countries started to increase. From 1945 onwards, the life expectancy for developing and under developed countries also started to increase as the World War 2 just ended.   
```{r}
plot(c(1800:2100), TopLifeMean, col="red", main = "Life Expentency in years", pch=1, ylim = c(0,150), xlab = "Years", ylab = "Life Expectency") 
points(c(1800:2100), MidLifeMean, col="blue")
points(c(1800:2100), LowLifeMean, pch=20, col="green")
abline(v= 1860)
abline(v= 1945)
text(1945, 100, "World War 2")
legend(x = "topright", legend = c("Developed countries ","Developing countries","Under Developed countries"), fill = c("red","blue","green"))
```
    
## DataSet: Total population    

**Finding the year wise mean for all the countries**    
```{r}
TopPopulationMean <- sapply(TopPopulation[,2:302], mean)
MidPopulationMean <- sapply(MidPopulation[,2:302], mean)
LowPopulationMean <- sapply(LowPopulation[,2:302], mean)
```


**Plotting the means for all the countries for all the 3 groups in the same plot, to compare the the trends**
```{r}
plot(c(1800:2100), TopPopulationMean, col="red", main = "Total Population", pch=1, ylim = c(1076000,63436000), xlab = "Years", ylab = "Popuation") 
points(c(1800:2100), MidPopulationMean, col="blue")
points(c(1800:2100), LowPopulationMean, pch=20, col="green")
legend(x = "topleft", legend = c("Developed countries ","Developing countries","Under Developed countries"), fill = c("red","blue","green"))
```
   
**Making box plot to compare the means over the course of years**
```{r}
boxplot(TopPopulationMean, MidPopulationMean, LowPopulationMean, ylab = "Population", names=c("Developed ","Developing ","Under Developed "))
```
   
**Analysis:**    
We see an exponential increase in the population of the under developed coountries in the plot. But the box plot shows us that there are many outliers and the average population is still considerably lower than that of the developed countries. This exponential increase might be due to poverty and lack of education awareness among the people in these countries. Before 1990, the developing and under developed countries had the similar trend in their population.    
```{r}
plot(c(1800:2100), TopPopulationMean, col="red", main = "Total Population", pch=1, ylim = c(1076000,63436000), xlab = "Years", ylab = "Population") 
points(c(1800:2100), MidPopulationMean, col="blue")
points(c(1800:2100), LowPopulationMean, pch=20, col="green")
abline(v = 1995)
legend(x = "topleft", legend = c("Developed countries ","Developing countries","Under Developed countries"), fill = c("red","blue","green"))
```
   
    
## DataSet: Income per person   
**Finding the year wise mean for all the countries**    
```{r}
TopIncomeMean <- sapply(TopIncome[,2:242], mean)
MidIncomeMean <- sapply(MidIncome[,2:242], mean)
LowIncomeMean <- sapply(LowIncome[,2:242], mean)
```


**Plotting the means for all the countries for all the 3 groups in the same plot, to compare the the trends**
```{r}
plot(c(1800:2040), TopIncomeMean, col="red", main = "GDP per capita", pch=1, ylim = c(0,80000), xlab = "Years", ylab = "GDP per capita") 
points(c(1800:2040), MidIncomeMean, col="blue")
points(c(1800:2040), LowIncomeMean, pch=20, col="green")
legend(x = "topleft", legend = c("Developed countries ","Developing countries","Under Developed countries"), fill = c("red","blue","green"))
```
   
**Making box plot to compare the means over the course of years**
```{r}
boxplot(TopIncomeMean, MidIncomeMean, LowIncomeMean, ylab = "GDP per capita", names=c("Developed ","Developing ","Under Developed "))
```
      
**Analysis:**  
We observe that the GDP dips during World War I and World War II   
```{r}
plot(c(1800:2040), TopIncomeMean, col="red", main = "GDP per capita", pch=1, ylim = c(0,80000), xlab = "Years", ylab = "GDP per capita") 
points(c(1800:2040), MidIncomeMean, col="blue")
points(c(1800:2040), LowIncomeMean, pch=20, col="green")
abline(v= 1918)
abline(v= 1945)
text(1918, 20075, "World War 1")
text(1945, 30075, "World War 2")
legend(x = "topleft", legend = c("Developed countries ","Developing countries","Under Developed countries"), fill = c("red","blue","green"))
```
   
   
## DataSet: Children per woman   
**Finding the year wise mean for all the countries**   
```{r}
TopChildrenMean <- sapply(TopChildren[,2:302], mean)
MidChildrenMean <- sapply(MidChildren[,2:302], mean)
LowChildrenMean <- sapply(LowChildren[,2:302], mean)
```

```{r}
plot(c(1800:2100), TopChildrenMean, col="red", main = "Children per woman", pch=1, ylim = c(0,10), xlab = "Years", ylab = "Number of children") 
points(c(1800:2100), MidChildrenMean, col="blue")
points(c(1800:2100), LowChildrenMean, pch=20, col="green")
legend(x = "topright", legend = c("Developed countries ","Developing countries","Under Developed countries"), fill = c("red","blue","green"))
```
   
**Making box plot to compare the means over the course of years**
```{r}
boxplot(TopChildrenMean, MidChildrenMean, LowChildrenMean, ylab = "Children per woman", names=c("Developed ","Developing ","Under Developed "))
```
   
**Analysis:** 
We see that on an average, under developed countries have more children per woman than developed and developing countries   
  
   
## DataSet: Child Mortality of 0-5 year olds dying per 1000 born    
    
**Finding the year wise mean for all the countries**   
```{r}
TopMortalityMean <- sapply(TopMortality[,2:302], mean)
MidMortalityMean <- sapply(MidMortality[,2:302], mean)
LowMortalityMean <- sapply(LowMortality[,2:302], mean)
```
    
**Plotting the means for all the countries for all the 3 groups in the same plot, to compare the the trends**
```{r}
plot(c(1800:2100), TopMortalityMean, col="red", main = "Child Mortality Rate", pch=1, ylim = c(0,500), xlab = "Years", ylab = "Child Mortality Rate") 
points(c(1800:2100), MidMortalityMean, col="blue")
points(c(1800:2100), LowMortalityMean, pch=20, col="green")
legend(x = "topright", legend = c("Developed countries ","Developing countries","Under Developed countries"), fill = c("red","blue","green"))
```
    
**Making box plot to compare the means over the course of years**
```{r}
boxplot(TopMortalityMean, MidMortalityMean, LowMortalityMean, ylab = "Child Mortality Rate", names=c("Developed ","Developing ","Under Developed "))
```
   
**Analysis:**   
The child mortality rate for all the countries have significantly gone down over the years.


   
## Now we plot all the datasets in a group wise manner and try to see the effect of one parameter in the others  
**For developed countries**   
We will normalize all the values using min-max normalization so that we can compare the values.

```{r}
normalize <- function(x) {
return ((x - min(x)) / (max(x) - min(x)))
}
```

```{r}
TopLifeMeanNorm <- normalize(TopLifeMean)
TopPopulationMeanNorm <- normalize(TopPopulationMean)
TopIncomeMeanNorm <- normalize(TopIncomeMean)
TopChildrenMeanNorm <- normalize(TopChildrenMean)
TopMortalityMeanNorm <- normalize(TopMortalityMean)
```


```{r}
plot(c(1800:2100), TopLifeMeanNorm, col="red", main = "Developed countries", pch=1, ylim = c(0,2), xlab = "Years", ylab = "Value") 
points(c(1800:2100), TopPopulationMeanNorm, col="blue")
points(c(1800:2040), TopIncomeMeanNorm, pch=20, col="green")
points(c(1800:2100), TopChildrenMeanNorm, pch=20, col="black")
points(c(1800:2100), TopMortalityMeanNorm, pch=20, col="orange")
legend(x = "topright", legend = c("Life Expectency ","Total Popuation","Income per person", "Children per woman", "Child mortality rate"), fill = c("red","blue","green", "black", "orange"))
```

    
**For developing countries**
      
```{r}
MidLifeMeanNorm <- normalize(MidLifeMean)
MidPopulationMeanNorm <- normalize(MidPopulationMean)
MidIncomeMeanNorm <- normalize(MidIncomeMean)
MidChildrenMeanNorm <- normalize(MidChildrenMean)
MidMortalityMeanNorm <- normalize(MidMortalityMean)
```

```{r}
plot(c(1800:2100), MidLifeMeanNorm, col="red", main = "Developing countries", pch=1, ylim = c(0,2), xlab = "Years", ylab = "Value") 
points(c(1800:2100), MidPopulationMeanNorm, col="blue")
points(c(1800:2040), MidIncomeMeanNorm, pch=20, col="green")
points(c(1800:2100), MidChildrenMeanNorm, pch=20, col="black")
points(c(1800:2100), MidMortalityMeanNorm, pch=20, col="orange")
legend(x = "topright", legend = c("Life Expectency ","Total Popuation","Income per person", "Children per woman", "Child mortality rate"), fill = c("red","blue","green", "black", "orange"))
```

**For under developed countries**

```{r}
LowLifeMeanNorm <- normalize(LowLifeMean)
LowPopulationMeanNorm <- normalize(LowPopulationMean)
LowIncomeMeanNorm <- normalize(LowIncomeMean)
LowChildrenMeanNorm <- normalize(LowChildrenMean)
LowMortalityMeanNorm <- normalize(LowMortalityMean)
```


```{r}
plot(c(1800:2100), LowLifeMeanNorm, col="red", main = "Under Developed countries", pch=1, ylim = c(0,2), xlab = "Years", ylab = "Value") 
points(c(1800:2100), LowPopulationMeanNorm, col="blue")
points(c(1800:2040), LowIncomeMeanNorm, pch=20, col="green")
points(c(1800:2100), LowChildrenMeanNorm, pch=20, col="black")
points(c(1800:2100), LowMortalityMeanNorm, pch=20, col="orange")
legend(x = "topright", legend = c("Life Expectency ","Total Popuation","Income per person", "Children per woman", "Child mortality rate"), fill = c("red","blue","green", "black", "orange"))
```
**Analysis:**    
In all the 3 groups of countries, we see that as the income per person goes up, the child mortality rate and children per woman decrease while life expectancy and total population increase.    
    
# Testing the correlation between income per person and the rest of the datasets  
## Income vs. Life expectancy
**For developed countries**
```{r}
cor.test(TopIncomeMean, TopLifeMean[1:241])
```
**For developing countries**
```{r}
cor.test(MidIncomeMean, MidLifeMean[1:241])
```
**For under developed countries**
```{r}
cor.test(LowIncomeMean, LowLifeMean[1:241])
```
**For all the categories, the correlation was close to 1, which means as the per person income increases, life expectancy increases.**   
   
## Income vs. Children per woman   
**For developed countries**
```{r}
cor.test(TopIncomeMean, TopChildrenMean[1:241])
```
**For developing countries**
```{r}
cor.test(MidIncomeMean, MidChildrenMean[1:241])
```
**For under developed countries**
```{r}
cor.test(LowIncomeMean, LowChildrenMean[1:241])
```
**For all the categories, the correlation was negative, which means as the per person income increases, children per woman decreases.**   
      
      
## Income vs. Child mortality rate   
**For developed countries**
```{r}
cor.test(TopIncomeMean, TopMortalityMean[1:241])
```
**For developing countries**
```{r}
cor.test(MidIncomeMean, MidMortalityMean[1:241])
```
**For under developed countries**
```{r}
cor.test(LowIncomeMean, LowMortalityMean[1:241])
```
**For all the categories, the correlation was negative, which means as the per person income increases, child mortality rate decreases.**   
      
      
## Income vs. Popuation   
**For developed countries**
```{r}
cor.test(TopIncomeMean, TopPopulationMean[1:241])
```
**For developing countries**
```{r}
cor.test(MidIncomeMean, MidPopulationMean[1:241])
```
**For under developed countries**
```{r}
cor.test(LowIncomeMean, LowPopulationMean[1:241])
```
**For all the categories, the correlation was close to 1, which means as the per person income increases, population increases linearly.**   
      
      
      
# Testing a few hypothesis on the datasets   

## Child Mortality Rate
**Shapiro-Wilk normality test**   
We first check if the data is normally distributed.   
-> Null hypothesis $H_0$ : The data is normally distributed     
-> Alternative hypothesis $H_d$ : The data is not normally distributed   
   
**For developed countries**
```{r}
shapiro.test(TopMortalityMean)
```
   
**For developing countries**
```{r}
shapiro.test(MidMortalityMean)
```
   
**For under developed countries**
```{r}
shapiro.test(LowMortalityMean)
```
For all the groups, the P-values are too low, suggesting that our alternate hypothesis is true, meaning that the data is not normally distrbuted.   
   
**Wilcoxon test**   
   
Now we will do a non parametric test to compare the means of our datasets pairwise. For that, we use the Wilcoxon test to compare the means.
   
**Case 1**    
-> Null hypothesis $H_0$ : The difference between the mean child mortality rate of developing and developed countries is zero     
-> Alternative hypothesis $H_d$ : The difference between the mean child mortality rate of developed and developing countries is less than zero i.e. developed countries are lower in child mortality rate.   
```{r}
wilcox.test(TopMortalityMean, MidMortalityMean, paired = TRUE, alternative = "less")
```
**Case 2**    
-> Null hypothesis $H_0$ : The difference between the mean child mortality rate of developed and under developed countries is zero     
-> Alternative hypothesis $H_d$ : The difference between the mean child mortality rate of developed and under developed countries is less than zero i.e. developed countries are lower in child mortality rate. 
```{r}
wilcox.test(TopMortalityMean, LowMortalityMean, paired = TRUE, alternative = "less")
```
**Case 3**    
-> Null hypothesis $H_0$ : The difference between the mean child mortality rate of developing and under developed countries is zero     
-> Alternative hypothesis $H_d$ : The difference between the mean child mortality rate of developing and under developed countries is less than zero i.e. developing countries are lower in child mortality rate.
```{r}
wilcox.test(MidMortalityMean, LowMortalityMean, paired = TRUE, alternative = "less")
```      
      
      
## Life Expectancy
**Shapiro-Wilk normality test**   
We first check if the data is normally distributed.   
-> Null hypothesis $H_0$ : The data is normally distributed     
-> Alternative hypothesis $H_d$ : The data is not normally distributed   
   
**For developed countries**
```{r}
shapiro.test(TopLifeMean)
```
  
**For developing countries**
```{r}
shapiro.test(MidLifeMean)
```
   
**For under developed countries**
```{r}
shapiro.test(LowLifeMean)
```
      
For all the groups, the P-values are too low, suggesting that our alternate hypothesis is true, meaning that the data is not normally distrbuted.   
   
**Wilcoxon test**   
   
Now we will do a non parametric test to compare the means of our datasets pairwise. For that, we use the Wilcoxon test to compare the means.
   
**Case 1**    
-> Null hypothesis $H_0$ : The difference between the mean life expectancy of developing and developed countries is zero      
-> Alternative hypothesis $H_d$ : The difference between the mean life expectancy of developed and developing countries is more than zero i.e. developed countries are higher in life expectancy.   
```{r}
wilcox.test(TopLifeMean, MidLifeMean, paired = TRUE, alternative = "greater")
```
**Case 2**    
-> Null hypothesis $H_0$ : The difference between the mean mean life expectancy of developed and under developed countries is zero      
-> Alternative hypothesis $H_d$ : The difference between the mean mean life expectancy of developed and under developed countries is more than zero i.e. developed countries are greater in mean life expectancy. 
```{r}
wilcox.test(TopMortalityMean, LowMortalityMean, paired = TRUE, alternative = "greater")
```
      
**Case 3**    
-> Null hypothesis $H_0$ : The difference between the mean life expectancy of developing and under developed countries is zero      
-> Alternative hypothesis $H_d$ : The difference between the mean life expectancy of developing and under developed countries is more than zero i.e. developing countries are greater in child mortality rate.
```{r}
wilcox.test(MidMortalityMean, LowMortalityMean, paired = TRUE, alternative = "greater")
```      
    
## Population
**Shapiro-Wilk normality test**   
We first check if the data is normally distributed.   
-> Null hypothesis $H_0$ : The data is normally distributed     
-> Alternative hypothesis $H_d$ : The data is not normally distributed   
   
**For developed countries**
```{r}
shapiro.test(TopPopulationMean)
```
  
**For developing countries**
```{r}
shapiro.test(MidPopulationMean)
```
   
**For under developed countries**
```{r}
shapiro.test(LowPopulationMean)
```
      
For all the groups, the P-values are too low, suggesting that our alternate hypothesis is true, meaning that the data is not normally distrbuted.   
   
**Wilcoxon test**   
   
Now we will do a non parametric test to compare the means of our datasets pairwise. For that, we use the Wilcoxon test to compare the means.
   
**Case 1**    
-> Null hypothesis $H_0$ : The difference between the mean population of developing and developed countries is zero   
-> Alternative hypothesis $H_d$ : The difference between the mean population of developed and developing countries is not zero.   
```{r}
wilcox.test(TopPopulationMean, MidPopulationMean, paired = TRUE)
```
**Case 2**    
-> Null hypothesis $H_0$ : The difference between the mean mean population of developed and under developed countries is zero     
-> Alternative hypothesis $H_d$ : The difference between the mean mean population of developed and under developed countries is not zero. 
```{r}
wilcox.test(TopPopulationMean, LowPopulationMean, paired = TRUE)
```
      
**Case 3**    
-> Null hypothesis $H_0$ : The difference between the mean population of developing and under developed countries is zero     
-> Alternative hypothesis $H_d$ : The difference between the mean population of developing and under developed countries is not zero.
```{r}
wilcox.test(MidPopulationMean, LowPopulationMean, paired = TRUE)
```      
    
   
## Children per woman
**Shapiro-Wilk normality test**   
We first check if the data is normally distributed.    
-> Null hypothesis $H_0$ : The data is normally distributed    
-> Alternative hypothesis $H_d$ : The data is not normally distributed   
   
**For developed countries**
```{r}
shapiro.test(TopChildrenMean)
```
  
**For developing countries**
```{r}
shapiro.test(MidChildrenMean)
```
   
**For under developed countries**
```{r}
shapiro.test(LowChildrenMean)
```
      
For all the groups, the P-values are too low, suggesting that our alternate hypothesis is true, meaning that the data is not normally distrbuted.   
   
**Wilcoxon test**   
   
Now we will do a non parametric test to compare the means of our datasets pairwise. For that, we use the Wilcoxon test to compare the means.
   
**Case 1**    
-> Null hypothesis $H_0$ : The difference between the mean children per woman of developing and developed countries is zero    
-> Alternative hypothesis $H_d$ : The difference between the mean children per woman of developed and developing countries is not zero.   
```{r}
wilcox.test(TopChildrenMean, MidChildrenMean, paired = TRUE)
```
**Case 2**    
-> Null hypothesis $H_0$ : The difference between the mean mean children per woman of developed and under developed countries is zero     
-> Alternative hypothesis $H_d$ : The difference between the mean mean children per woman of developed and under developed countries is not zero. 
```{r}
wilcox.test(TopChildrenMean, LowChildrenMean, paired = TRUE)
```
      
**Case 3**    
-> Null hypothesis $H_0$ : The difference between the mean children per woman of developing and under developed countries is zero     
-> Alternative hypothesis $H_d$ : The difference between the mean children per woman of developing and under developed countries is not zero.
```{r}
wilcox.test(MidChildrenMean, LowChildrenMean, paired = TRUE)
```      

**Note:**    
For each category of countries, the shapiro test told us that the datas aren't normally distributed. Hence we cannot use **t test or ANOVA** to compare the means of our datasets.